---
title: "Basic Inferential Data Analysis of ToothGrowth Dataset"
author: "Jeevithiesh Duggani"
date: "07/05/2020"
output: 
    pdf_document:
        toc: true
        toc_depth: 3
---
```{r Set global Settings,eval=TRUE,results='hide',echo=FALSE}
    options(scipen = 100)
```

## Abstract
This project analyzes the `ToothGrowth` data in the R datasets package.

The `aims` of the project are listed below :

1. Load the `ToothGrowth` data and perform some `basic exploratory data analyses`.
2. Provide a `basic summary` of the data.
3. Use `confidence intervals` and/or `hypothesis tests` to compare `tooth growth` by **supp** and **dose**.
4. State the `conclusions` and the `assumptions` needed for the **conclusions**.

## Import the necessary packages and datasets
The following packages are to be imported.

* dataset
* ggplot2
```{r Import Packages,results='hide'}
    library(datasets)
    library(ggplot2)
```

Import the following datasets.

* ToothGrowth
```{r Import Datasets,results='hide'}
    data("ToothGrowth")
```

Rename the values `OJ` as `Orange Juice` and `VC` as `Ascorbic Acid`.
```{r Rename Values}
    levels(ToothGrowth$supp) <- c("Orange Juice", "Ascorbic Acid")
```

## Run Basic Exploratory Data Analyses
Check the `structure` of the dataset `ToothGrowth`.
```{r Structure}
    str(ToothGrowth)
```

supp consists of **2** `unique` values. Identify number of unique values in `len` and `dose`.
```{r Identify unique values}
    list(uniqLen  = length(unique(ToothGrowth$len)), 
         uniqDose = length(unique(ToothGrowth$dose)))
```

Convert dose into a factor as it consists of only **3** unique values.
```{r Dose as factor variable}
    ToothGrowth$dose <- as.factor(ToothGrowth$dose)
    str(ToothGrowth)
```

Plot a boxplot to visualize the `length of the teeth` based on the `supplement` and `dosage`.
```{r Graph length vs dose}
    graph <- ggplot(ToothGrowth, aes(dose,len))
    graph <- graph + facet_grid(.~supp)
    graph <- graph + geom_boxplot(aes(fill = supp), show.legend = FALSE)
    graph <- graph + labs(title = "Tooth growth by Dosage for each Supplement", 
                  y = "Length of the tooth", x = "Dosage (mg / day)")
    graph
```

## Observations

* It is observed that **both** `supplements` result in an **increase** in `growth` as the `dosage` is **increased**.
* Both supplements result in the **same** `length` when the `dosage` is at `2 mg / day`.
* `Orange Juice` is **more** `effective` at the **lower** `doses`.

## Comparision of length of growth by supplement and dose
Use `Confidence Intervals` and `Hypothesis Testing` for the comparision.

### Hypothesis 1
`Orange Juice` and `Ascorbic Acid` result in the same growth.
```{r Testing Hypothesis 1}
    test1 <- t.test(len ~ supp, data = ToothGrowth)
    conf1 <- test1$conf.int
    pval1 <- test1$p.value
    test1
```

- The confidence interval is `r conf1` which contains the value `0`.
- The P value is `r pval1` which is **greater** than `0.05`.

> Therefore the null hypothesis cannot be rejected

### Hypothesis 2
`Orange Juice` and `Ascorbic Acid` result in the same growth at a dosage of 0.5 mg / day.
```{r Testing Hypothesis 2}
    test2 <- t.test(len ~ supp, data = ToothGrowth[ToothGrowth$dose == 0.5,])
    conf2 <- test2$conf.int
    pval2 <- test2$p.value
    test2
```

- The confidence interval is `r conf1` which does not contains the value `0`.
- The P value is `r pval1` which is **much lesser** than `0.05`.

> Therefore the null hypothesis can be rejected. The alternative hypothesis "The supplement Orange Juice results in more tooth growth at a dosage of 0.5 mg / day" is accepted instead.

### Hypothesis 3
`Orange Juice` and `Ascorbic Acid` result in the same growth at a dosage of 1 mg / day.
```{r Testing Hypothesis 3}
    test3 <- t.test(len ~ supp, data = ToothGrowth[ToothGrowth$dose == 1,])
    conf3 <- test3$conf.int
    pval3 <- test3$p.value
    test3
```

- The confidence interval is `r conf3` which does not contains the value `0`.
- The P value is `r pval3` which is **much smaller** than `0.05`.

> Therefore the null hypothesis can be rejected. The alternative hypothesis "The supplement Orange Juice results in more tooth growth at a dosage of 1 mg / day" is accepted instead.

### Hypothesis 4
`Orange Juice` and `Ascorbic Acid` result in the same growth at a dosage of 2 mg / day.
```{r Testing Hypothesis 4}
    test4 <- t.test(len ~ supp, data = ToothGrowth[ToothGrowth$dose == 2,])
    conf4 <- test4$conf.int
    pval4 <- test4$p.value
    test4
```

- The confidence interval is `r conf4` which contains the value `0`.
- The P value is `r pval4` which is **much greater** than `0.05`.

> Therefore the null hypothesis cannot be rejected

## Assumptions

- `Normal Disibution` of the lengths of the tooth.
- `No other measured factors` affects the length of the tooth.
- For the `t tests` the variances of the two groups are different from one another.

## Conclusions

- `Orange Juice` results in more growth at the **lower** doses of `0.5 and 1 mg / day`.
- `Orange Juice` and `Ascorbic Acid` results in the same growth at the **highest** doses of `2 mg / day`.
- For the overall picture, We cannot conclude that `Orange Juice` is more effective than `Ascorbic Acid`.

                                **End**