---
title: "Investigation of the exponential distribution in R"
author: "Jeevithiesh Duggani"
date: "06/05/2020"
output: 
    pdf_document:
        toc: true
---

```{r Set global Settings,eval=TRUE,results='hide',echo=FALSE}
    options(scipen = 100)
```

## Abstract
This project investigates the `exponential distribution` in R and compare it with the `Central Limit Theorem`. It investigates the distribution of **averages** of **forty exponentials**. **Thousand** simulations are performed.

The `aims` of the project are listed below :

1. Show the `sample mean` and compare it to the `theoretical mean` of the distribution.
2. Show how variable the sample is (via `variance`) and compare it to the `theoretical variance` of the distribution.
3. Show that the distribution is approximately normal, i.e. the difference between the distribution of a `large collection of random exponentials` and the distribution of a `large collection of averages of 40 exponentials` is **negligible**.

## Import the necessary packages
The following packages are to be imported.

* ggplot2
```{r Import Packages,results='hide'}
    library(ggplot2)
```

## Initialization of important Variables
The `seed` is initialized with `100`.
The `lambda (rate parameter)` is set as `0.2`.
The `sample size` is set to `40`.
The `number of simulations` is `1000`.

```{r Initialization of Important Variables}
    set.seed(100)

    lambda      <- 0.2
    nsim        <- 1000
    sampleSize  <- 40
```

## Run the simulations
The simulations are run and stored in the matrix `sims`.
```{r Generate nsim simulations and store them in sims}
    sims <- matrix(rexp(nsim*sampleSize, lambda), nsim, sampleSize)
```

## Comparision between Sample Mean and Theoritical Mean
The value of the Theoritical mean is `1/lambda = 5`.

The Sample mean is calculated below.
```{r Calculate the Theoritical Mean and the Sample Mean}
    tMean <- 1/lambda # Theoritical Mean
    sMean <- mean(rowMeans(sims)) #Sample Mean
```

The value of the Sample mean is `r round(sMean,6)`.

The difference between the two means is `r round(sMean - tMean,6)`, which is `r paste0(round((sMean - tMean)/tMean,6),"%")` of the Theoritical mean.

## Comparision between Sample Variance and Theoritical Variance
The value of the Theoritical variance is `(1/lambda^2)*(1/sampleSize) = 0.625`.

The Sample mean is calculated below.
```{r Calculate the Theoritical Variance and the Sample Variance}
    tVar <- (1/lambda^2)*(1/sampleSize) # Theoritical Variance
    sVar <- var(rowMeans(sims)) #Sample Variance
```

The value of the Sample variance is `r round(sVar,6)`.

The difference between the two variances is `r round(sVar - tVar, 6)`, which is `r paste0(round((sVar - tVar)/tVar, 6),"%")` of the Theoritical variance.

## Comparisions of the Distributions
Plot the Theoritical as well as the sample Distributions.

```{r Plot the graphs for the Sample and the Theoritical Distribution}
    graph <- ggplot(data.frame(mean = rowMeans(sims)), aes(mean))
    graph <- graph + geom_histogram(aes(y=..density..), binwidth = 0.2, 
                                    fill = "wheat", color = "black")
    graph <- graph + stat_function(fun = dnorm, args = list(mean = tMean, sd = sqrt(tVar)))
    graph <- graph + labs(title = "Theoritical and Sample Simulations", 
                          x = "Mean", y = "Probability Density")
    graph
```

Both the distributions follow the same pattern.

## Results

1. Theoritical Mean is `r tMean`, Sample Mean is `r round(sMean,6)`. The difference between the two means is `r round(sMean - tMean, 6)`, which is `r paste0(round((sMean - tMean)/tMean, 6),"%")` of the Theoritical Mean.

2. Theoritical Variance is `r tVar`, Sample Variance is `r round(sVar, 6)`. The difference between the two variances is `r round(sVar - tVar, 6)`, which is `r paste0(round((sVar - tVar)/tVar,6),"%")` of the Theoritical Variance.

3. Both the distributions follow the same pattern.

                                **End**